Ion pairing controls rheological properties of polyelectrolyte crocodile line hydrogels

### 2016-01-11 PALSE/ion_pairing_paper.tex
Add the sketch of the network in fig:dissociation and update the text accordingly

\begin{scheme}
\includegraphics{metathesis.pdf}
\caption{(a) Anionic metathesis reaction. As above, \ce{Nu+} corresponds either to pyrrolidinium or imidazolium. (b) Summary of the polymers derived from \ce{PBr}.}
\label{sch:metathesis}
\end{scheme}

@@

\subsection{Crocodile chain model}
We have previously demonstrated that, due to the opposite charge of the phosphonate head, head-to-body ionic bonds are possible~\cite{Srour2014}. If at least two foreign heads attach to the same body, we obtain an effective cross-linking point. When the probability of such a configuration is non-zero, we obtain a physically cross-linked gel.

@@

We thus have two limiting cases: (i) every chain has at least two heads attached and we have roughly $N_0$ monomers between cross-linking points; (ii) every chain has a single head attached and we have a single supramolecular chain in the system. The former case should be observed at high charge dissociation when many moieties are ionised along the chain providing many sites for a phosphonate head to bound to. The latter case should be observed at negligible charge dissociation. In between we should observe cross-linking points separated by crocodile lines of $n N_0$ monomers.

@@

\begin{table}
\begin{tabular}{l|SS|SSS}
& {$G^\prime(\gamma\rightarrow 0)$} & {$\gamma_c$} & {$n$} & {$E_c$} & {$A$}\\
&	{\si{\pascal}} &  & & {\si{\kilo\joule/\mol}} & \\\hline&&&&\\[-10pt]
\ce{PPyr+I-}	& 8800	&	0.04	&	1.	&	0.004\cellcolor{gray!25}	&	1.\\
\ce{PPyr+Br-}	& 9800 	& 	0.016	&	1.05&	0.001\cellcolor{gray!25}	&	1.1\\
\ce{PPyr+Cl-}	& 270 	&	0.09	&	45	&	0.02\cellcolor{gray!25}	&	7.\\
\ce{PIm+I-}	& 78	&	0.16	&	113	&	0.063\cellcolor{gray!25}	&	9.6\\
\ce{PIm+Br-}	& 13	&	8.04	&	790	&	160.	&	3.4\cellcolor{gray!25}\\
\end{tabular}
\caption{Summary of rheological measurements and microscopic values deduced from the model. Gray background indicates incoherent values obtained either by (Eq.~\ref{eq:Ec}) or by (Eq.~\ref{eq:A}) outside of their respective validity domain.}
\label{tab:results}
\end{table}

@@

@@

%\begin{figure}
%\begin{center}
%\begin{tikzpicture}
%\begin{axis}[
%	scale only axis,
%	width=12\baselineskip, height=6\baselineskip, 
%	axis lines=none,
%	xmin=0.5, xmax=4.5,
%	ymin=-0.7, ymax=1.3,
%	]
%\addplot[only marks, mark=o, mark size=2] table {/home/mathieu/tex/PALSE/persistent.xy} coordinate[pos=0.43] (a) coordinate[pos=0.57] (b) coordinate[pos=0.81](c);
%\draw[<->] ($(a)+(-0.5em,0)$) -- ($(b)+(-0.5em,0)$) node[midway, left]{$r_\mathrm{scr}$};
%\draw[<->] ($(c)+(-0.25em,-0.5em)$) -- ($(c)+(0.25em,-0.5em)$) node[midway, below]{$D$};
%\end{axis}
%\end{tikzpicture}
%\end{center}
%\caption{Sketch of the persistent random walk.}
%\label{fig:persistent}
%\end{figure}

@@

The spatial conformation of a polyelectrolyte chain in semi-dilute, low salt solution is a persistent random walk where the persistence length is the screening length $r_\mathrm{src}$ due to the dissociated counter ions~\cite{Rubinstein1996}. In bad solvent, the persistent random walk is made of electrostatic blobs of diameter $D$ containing $g_e$ monomers, see Figure~\ref{fig:persistent}:
\begin{align}
g_e &= \frac{A^2}{u}\tau\label{eq:ge},\\
D &= b \left(\frac{A^2}{u}\right)^{1/3}\label{eq:D}.
\end{align}
Here $A$ is the number of monomers between uncondensed charges, $\tau = 1 - T/\Theta$ the reduced temperature and $u = \ell_\mathrm{B}/b \approx 2.7$ is the ratio of Bjerrum length $\ell_\mathrm{B}$ and monomer size $b\approx \SI{0.26}{\nano\metre}$.

@@

\begin{figure}
\begin{center}
\includegraphics{dissociation.pdf}
\end{center}
\caption{Relation between the number of monomers between uncondensed charges and the number of chains between cross-link points. The line corresponds to $n = A^2$.}
\label{fig:dissociation}
\end{figure}

>>>>

\begin{scheme}
\includegraphics{metathesis.pdf}
\caption{(a) Anionic metathesis reaction. As above, \ce{Nu+} corresponds either to pyrrolidinium or imidazolium. (b) Summary of the polymers derived from a single batch of \ce{PBr}.}
\label{sch:metathesis}
\end{scheme}

@@

\subsection{Crocodile chain model}

\begin{figure}
\includegraphics{dissociation.pdf}
\caption{Crocodile-line model. (a) Sketch of the network in the case of a crocodile line of size $n=3$ between cross-links (gray disks) and a persistence length $r_\mathrm{src}\approx 3D$. Empty circles are not individual monomers but electrostatic blobs of size $D$. Electrostatic blobs containing an anionic head are shown as black filled circles. (b) Relation between the number of monomers between uncondensed charges and the number of chains between cross-link points. The line corresponds to $n = A^2$.}
\label{fig:dissociation}
\end{figure}

We have previously demonstrated that, due to the opposite charge of the phosphonate head, head-to-body ionic bonds are possible~\cite{Srour2014}. If at least two foreign heads attach to the same body, we obtain an effective cross-linking point. When the probability of such a configuration is non-zero, we obtain a physically cross-linked gel.

@@

We thus have two limiting cases: (i) every chain has at least two heads attached and we have roughly $N_0$ monomers between cross-linking points; (ii) every chain has a single head attached and we have a single supramolecular chain in the system. The former case should be observed at high charge dissociation when many moieties are ionised along the chain providing many sites for a phosphonate head to bound to. The latter case should be observed at negligible charge dissociation. In between we should observe cross-linking points separated by crocodile lines of $n N_0$ monomers as sketched in Figure~\ref{fig:dissociation}a.

@@

@@

\begin{table}
\begin{tabular}{l|SS|SSSS}
& {$G^\prime(\gamma\rightarrow 0)$} & {$\gamma_c$} & {$n$} & \multicolumn{2}{c}{$E_c$} & {$A$}\\
&	{\si{\pascal}} &  & & {$k_\mathrm{B}T$} & {\si{\kilo\joule/\mol}} & \\\hline&&&&\\[-10pt]
\ce{PPyr+I-}	& 8800	&	0.04	&	1.	&	2e-3\cellcolor{gray!25}	&	0.004\cellcolor{gray!25}	&	1.\\
\ce{PPyr+Br-}	& 9800 	& 	0.016	&	1.05&	3e-4\cellcolor{gray!25}	&	0.001\cellcolor{gray!25}	&	1.1\\
\ce{PPyr+Cl-}	& 270 	&	0.09	&	45	&	8e-3\cellcolor{gray!25}	&	0.02\cellcolor{gray!25}	&	7.\\
\ce{PIm+I-}	& 78	&	0.16	&	113	&	3e-2\cellcolor{gray!25}	&	0.063\cellcolor{gray!25}	&	9.6\\
\ce{PIm+Br-}	& 13	&	8.04	&	790	&	64	&	160.	&	3.4\cellcolor{gray!25}\\
\end{tabular}
\caption{Summary of rheological measurements and microscopic values deduced from the model. Gray background indicates incoherent values obtained either by (Eq.~\ref{eq:Ec}) or by (Eq.~\ref{eq:A}) outside of their respective validity domain.}
\label{tab:results}
\end{table}

@@

@@

The spatial conformation of a polyelectrolyte chain in semi-dilute, low salt solution is a persistent random walk where the persistence length is the screening length $r_\mathrm{src}$ due to the dissociated counter ions~\cite{Rubinstein1996}. In bad solvent, the persistent random walk is made of electrostatic blobs of diameter $D$ containing $g_e$ monomers, see Figure~\ref{fig:dissociation}a:
\begin{align}
g_e &= \frac{A^2}{u}\tau\label{eq:ge},\\
D &= b \left(\frac{A^2}{u}\right)^{1/3}\label{eq:D}.
\end{align}
Here $A$ is the number of monomers between uncondensed charges, $\tau = 1 - T/\Theta$ the reduced temperature and $u = \ell_\mathrm{B}/b \approx 2.7$ is the ratio of Bjerrum length $\ell_\mathrm{B}$ and monomer size $b\approx \SI{0.26}{\nano\metre}$.

@@
